---
title: "ada-week-2.0"
author: "Ingrid Lundeen"
date: "September 15, 2016"
output: html_document
---

m<-matrix(data=c(1,2,3,4),nrow=2,ncol=2) 
m m<-matrix(data=c(1,2,3,4,5,6),nrow=2,ncol=3,byrow=TRUE) 
m V1<-c(1,2,3,4) 
V2<-c(6,7,8,9) 
m1<-rbind(V1,V2) 
m1 
m2<-cbind(V1,V2) 
m2 
class(m1) 
class(m2) 
class(V1) 
dim(m1) 
dim(m2) 
colnames(m2) 
m2[,2] 
rownames(m2) 
str(m2) 
#an array can have as many dimensions as you want - much more general of a structure 
#here the data is a set of numbers from 1 to 90 divided into 3 dimensions 
#here 5 is the number of rows per block, 6 is the number of columns per block, and 3 is the number of blocks 
a<-array(data=1:90,dim=c(5,6,3)) 
a v<-1:100 
v 
#using brackets to get out particular subsets of information from a matrix/array/etc 
#you are asking for the vector from index position 1 to 15 
v[1:15] 
#you can also use a vector to pull out specific things 
v[c(2,4,6,8,10)] 
#you can also use a function or calc to subset a vector 
v<-101:200 
v[seq(from=1,to=100,by=2)]


library(stringr)
```{r}
###CHALLENGE 1: create vector of first lines of gettysburg address + extract every third element of this vector
library(stringr)

ga<-"Four score and seven years ago our fathers brought forth on this continent a new nation conceived in Liberty and dedicated to the proposition that all men are created equal"
split<-str_split(ga," ", simplify=TRUE)
split

option1<-matrix(data=split,ncol=3,nrow=10,byrow=TRUE) 
option1
option1[,1]

option2<-split[seq(from=1, to=30, by=3)]
option2
```
```{r}
###Challenge 2
m<-matrix(data=1:80,nrow=8,ncol=10,byrow=FALSE)
m
```
```{r}
##Given the matrix, m, above, extract the 2nd, 3rd, and 6th columns and assign them to the variable x
two<-m[,2]
three<-m[,3]
six<-m[,6]
x<-cbind(two,three,six)
x
```

```{r}
##Given the matrix, m, above, extract the 6th to 8th row and assign them to the variable x
x<-m[6:8,]
x
```

```{r}

##Given the matrix, m, above, extract the elements from row 2, column 2 to row 6, column 9 and assign them to the variable x
z<-m[,2:9]
z

x<-z[2:6,]
x
```

```{r}
###Challenge3 : Construct a 4-dimensional, 400 element array (5 x 5 x 4 x 4) named a consisting of the numbers 400 to 1 (i.e., a descending series)
a<-array(data=400:1,dim=c(5,5,4,4))
a
```

```{r}
#Given this matrix, what would the following return?
#a[1, 1, 1, 2]???
a[1,1,1,2]

#a[2, 3, 2, ]???
a[2, 3, 2, ]

#a[1:5, 1:5, 3, 3]??
a[1:5, 1:5, 3, 3]
```

```{r}
###CHALLENGE 4: simplified list of primate taxonomy 

Anthropoidea<-list(Platyrrhini=c("Cebidae","Atelidae","Pitheciidae"),Catarrhini=c("Cercopithecidae","Hylobatidae","Hominidae"))
Anthropoidea

Tarsioidea<-c("Tarsiidae")
Haplorhini<-list(Anthropoidea, Tarsioidea)
names(Haplorhini)<-c("Anthropoidea","Tarsioidea")
Strepsirrhini<-list(Lorisoidea=c("Lorisidae","Galagidae"),Lemuroidea=c("Cheirogaleidae","Lepilemuridae","Indriidae","Lemuridae","Daubentoniidae"))
Primates<-list(Strepsirrhini=c(Strepsirrhini),Haplorhini=c(Haplorhini))
Primates
```

```{r}
###CHALLENGE 5: Store the following matrix: 3 0 1 23 1 2 33 1 1 42 0 1 41 0 2
t<-matrix(data=c(3,0,1,23,1,2,33,1,1,42,0,1,41,0,2),nrow=5,ncol=3)
t
```

```{r}
#Coerce the matrix to a data frame.
df_t<-as.data.frame(t)
df_t
```

```{r}
#As a data frame, coerce the second column to be logical-valued
df_t$V2<-as.logical(df_t$V2)
```

```{r}
#As a data frame, coerce the third column to be factor-valued
df_t$V3<-as.factor(df_t$V3)
df_t
```


?list to create a list use the list operator list()

f<-file.choose() 
df<-read.csv(f,sep=“,”,header=TRUE,stringsAsFactors=FALSE) 
df 
str(df) 
df[,1] str(df[,1]) dflastNamestr(df[1])df["lastName"]str(df["lastName"])dflastNamestr(df[1])df["lastName"]str(df["lastName"])dfschool <- c(“COLA”, “COLA”, “COLA”, “CNS”, “CNS”, “COLA”, “COLA”, “iSchool”, “CNS”, “CNS”, “COLA”, “iSchool”, “COLA”) df new_df<-df[df$school==“COLA”, ] new_df new_df<-df[df$school==“COLA”,c(“lastName”,“firstName”,“school”)] new_df library(evaluate) update.packages(evaluate)
